---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prqlr

<!-- badges: start -->
[![prqlr status badge](https://eitsupi.r-universe.dev/badges/prqlr)](https://eitsupi.r-universe.dev)
[![CRAN status](https://www.r-pkg.org/badges/version/prqlr)](https://CRAN.R-project.org/package=prqlr)
<!-- badges: end -->

R bindings for [the `prql-compiler` Rust library](https://github.com/prql/prql),
powered by [the `extendr` framework](https://extendr.github.io/).

## Installation

Requires R 4.2.0 or later.

This package can be installed from CRAN or [R-universe](https://eitsupi.r-universe.dev/ui#package:prqlr).
If available, a binary package will be installed.

```r
# Install from CRAN
install.packages("prqlr")
```

```r
# Install from R-universe
install.packages("prqlr", repos = "https://eitsupi.r-universe.dev")
```

For source installation, the Rust toolchain must be configured.
Please check the <https://github.com/r-rust/hellorust> repository.

## Example

```{r}
library(prqlr)
"from mtcars | filter cyl > 6 | select [cyl, mpg]" |>
  prql_compile() |>
  cat()
```

PRQL's pipelines can be joined by the newline character (`\n`), or actual newlines in addition to `|`.

```{r}
"from mtcars \n filter cyl > 6 \n select [cyl, mpg]" |>
  prql_compile() |>
  cat()
```

```{r}
"from mtcars
filter cyl > 6
select [cyl, mpg]" |>
  prql_compile() |>
  cat()
```

Thanks to the `{tidyquery}` package,
we can even convert a PRQL query to a SQL query and then to a `{dplyr}` query!

```{r}
"from mtcars
filter cyl > 6
select [cyl, mpg]" |>
  prql_compile() |>
  tidyquery::show_dplyr()
```

## Development

Use the `rextendr::document()` function to generate R source code from Rust source code in the `src` directory.
