---
title: Use PRQL with knitr
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use PRQL with knitr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Loading `{prqlr}` makes the `prql` code chunks executable in R Markdown and Quarto Markdown files.

## Output formats

It can either output the results of the actual query execution or display a SQL query compiled from a PRQL query.

### Use with `{DBI}` connections

If using with database connections (set the name of `{DBI}` connection to the `connection` chunk option),
PRQL code chunks pass the output SQL to `{knitr}`'s SQL engine and
behaves like [SQL code chunks](https://bookdown.org/yihui/rmarkdown/language-engines.html#sql).
So, usage is the same as for SQL code chunks.

For example, let's render an R Markdown file named `test.Rmd` with the following contents
with the `knitr::knit()` function.

````{r}
#| echo: false
#| results: asis

# Override the default SQL print function to avoid inserting '<div class="knitsql-table">'
# https://github.com/yihui/xaringan/issues/307
knitr::opts_knit$set(sql.print = \(x) paste(knitr::kable(x, format = "markdown"), collapse = "\n"))

src_with_con <- r"(
```{r}
#| echo: false

library(DBI)
library(prqlr)
con <- dbConnect(RSQLite::SQLite(), ":memory:")
dbWriteTable(con, "mtcars", mtcars)
```

```{prql}
#| connection: con

from mtcars
filter cyl > 6
select [cyl, mpg]
derive [mpg_int = round 0 mpg]
take 3
```
)"

cat("````markdown", src_with_con, "````", sep = "")
````

After rendering, a Markdown file named `test.md` is generated with the following contents.

````{r}
#| echo: false
#| results: asis

res_with_con <- knitr::knit_child(text = src_with_con, quiet = TRUE)
cat("````markdown", res_with_con, "````", sep = "")
````

Note that the syntax highlighting of Elm is the best for PRQL,
so the output code blocks are marked as `elm`.
This can be changed by setting the `lang` chunk option.

### Without database connections

PRQL code chunks without the `connection` option will generate SQL code blocks.

````{r}
#| echo: false
#| results: asis

src_without_con <- r"(
```{r}
#| echo: false

library(prqlr)
```

```{prql}
from mtcars
filter cyl > 6
select [cyl, mpg]
derive [mpg_int = round 0 mpg]
take 3
```
)"

cat("````markdown", src_without_con, "````", sep = "")
````

The R Markdown file above will be converted to the Markdown file below.

````{r}
#| echo: false
#| results: asis

res_without_con <- knitr::knit_child(text = src_without_con, quiet = TRUE)
cat("````markdown", res_without_con, "````", sep = "")
````

## Engine options

We can pass some options of `prql_compile()` via the chunk option `engine-opts` (or `engine.opts`).

Note that the `format` option is always `TRUE` for PRQL code chunks.

````{r}
#| echo: false
#| results: asis

src_engine_opts <- r"(
```{r}
#| echo: false

library(prqlr)
```

## YAML-style

```{prql}
#| engine-opts:
#|   target: sql.mssql
#|   signature_comment: false

from mtcars
take 3
```

## R-style

```{prql engine.opts=list(target="sql.mssql", signature_comment=FALSE)}
from mtcars
take 3
```
)"

# Check if it can be rendered
invisible(knitr::knit_child(text = src_engine_opts, quiet = TRUE))

cat("````markdown", src_engine_opts, "````", sep = "")
````
