---
title: Use PRQL on R
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use PRQL on R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package provides simple function to convert [PRQL](https://prql-lang.org/) query strings to SQL strings.

```{r}
library(prqlr)
"from mtcars | filter cyl > 6 | select [cyl, mpg]" |>
  prql_to_sql()
```

It can be combined with already existing great packages that manipulate data with SQL.

## Work with DB

Using it with the `{DBI}` package, we can execute PRQL queries against the database.

```{r}
library(DBI)

# Create an ephemeral in-memory RSQLite database
con <- dbConnect(RSQLite::SQLite(), ":memory:")

dbWriteTable(con, "mtcars_table", mtcars)

"from mtcars_table
filter cyl == 4" |>
  prql_to_sql() |>
  dbGetQuery(con, statement = _)
```

## Work with R data.frame

Using it with the `{tidyquery}` package, we can execute PRQL queries against R `data.frame`.

```{r}
library(tidyquery)
library(nycflights13)

"from flights
join side:left planes [tailnum]
filter (distance | in 200..300)
filter null != air_time
group [origin, dest] (
  aggregate [
    num_flts = count,
    num_seats = (sum seats | round 0),
    avg_delay = (average arr_delay | round 0)
  ]
)
filter num_flts > 3000
sort [-num_seats, avg_delay]
take 2
" |>
  prql_to_sql() |>
  query()
```
