---
title: Use PRQL on R
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use PRQL on R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[PRQL](https://prql-lang.org/) (Pipelined Relational Query Language, pronounced "Prequel")
is a modern language for transforming data, can be compiled to SQL.

This package provides a simple function to convert a PRQL query string to a SQL string.

For example, the following PRQL query can be converted to the below SQL query.

```elm
# This is a PRQL query
from mtcars
filter cyl > 6
select [cyl, mpg]
derive [mpg_int = round 0 mpg]
```

```sql
-- This is a SQL query
SELECT
  cyl,
  mpg,
  ROUND(mpg, 0) AS mpg_int
FROM
  mtcars
WHERE
  cyl > 6
```

Just pass the PRQL string to the `prql_to_sql()` function like this.

```{r}
library(prqlr)

"
from mtcars
filter cyl > 6
select [cyl, mpg]
derive [mpg_int = round 0 mpg]
" |>
  prql_to_sql()
```

This outputed SQL string can be used with already existing great packages that manipulate data with SQL.

## Work with DB

Using it with the `{DBI}` package, we can execute PRQL queries against the database.

Here sqlite is used via `{RSQLite}`,
so the [dialect](https://prql-lang.org/book/queries/dialect_and_version.html) header
is set at the top of the query.

```{r}
library(DBI)

# Create an ephemeral in-memory RSQLite database
con <- dbConnect(RSQLite::SQLite(), ":memory:")

dbWriteTable(con, "mtcars_table", mtcars)

"
prql dialect:sqlite

from mtcars_table
filter cyl > 6
select [cyl, mpg]
derive [mpg_int = round 0 mpg]
" |>
  prql_to_sql() |>
  dbGetQuery(con, statement = _)
```

## Work with R data.frame

Using it with the `{tidyquery}` package, we can execute PRQL queries against R `?data.frame`.

Let's run a query that joins and aggregates two data.frames, `flights` and `planes`,
contained in the `{nycflights13}` package.

```{r}
library(tidyquery)
library(nycflights13)

"
from flights
join side:left planes [tailnum]
filter (distance | in 200..300)
filter air_time != null
group [origin, dest] (
  aggregate [
    num_flts = count,
    num_seats = (sum seats | round 0),
    avg_delay = (average arr_delay | round 0)
  ]
)
filter num_flts > 3000
sort [-num_seats, avg_delay]
take 2
" |>
  prql_to_sql() |>
  query()
```

Note that [the PRQL syntax](https://prql-lang.org/book/) is very similar to the `{dplyr}` syntax.
`{dplyr}` is a very popular R package for manipulating data.

```{r}
library(dplyr, warn.conflicts = FALSE)
library(nycflights13)

flights |>
  left_join(planes, by = "tailnum") |>
  filter(
    distance |> between(200, 300),
    !is.na(air_time)
  ) |>
  group_by(origin, dest) |>
  summarise(
    num_flts = n(),
    num_seats = sum(seats, na.rm = TRUE) |> round(0),
    avg_delay = mean(arr_delay, na.rm = TRUE) |> round(0),
    .groups = "drop"
  ) |>
  filter(num_flts > 3000) |>
  arrange(desc(num_seats), avg_delay) |>
  head(2)
```
